---
title: 'Homework 4: Logistic Regression and kNN'
output: html_notebook
---
####Casey Erbes

#####This rmd script is an exercise in running logistic regression and applying the kNN algorithm in R.

##1

```{r}
# setup in console:
# install.packages('ISLR')
#
# a
library(ISLR)
# b
medianMpg <- median(Auto$mpg)
cat("Median value for mpg is:", medianMpg, "\n\n")
# c
Auto$mpglevel <- ifelse(Auto$mpg > medianMpg, 1, 0)
cat("Is mpglevel a factor?", is.factor(Auto$mpglevel), "\n\n")
Auto$mpglevel <- as.factor(Auto$mpglevel)
cat("Is mpglevel a factor now?", is.factor(Auto$mpglevel), "\n\n")
# d
names(Auto)
```

##2

```{r}
attach(Auto)
# a
par(mfrow=c(2,2))
# b
# plot point is green if mpglevel is 1, red otherwise.
plot(horsepower~mpg, col=ifelse(mpglevel==1,'green','red'))
plot(weight~mpg, col=ifelse(mpglevel==1,'green','red'))
# c
plot(horsepower~mpglevel)
plot(weight~mpglevel)
detach(Auto)
```

##3

```{r}
# split into train and test sets
set.seed(1234)
sampleInt <- sample.int(n = nrow(Auto), size = floor(.75*nrow(Auto)))
# exclude name and mpg in columns 1 and 9 respectively
train <- Auto[sampleInt, c(-1, -9)]
test  <- Auto[-sampleInt, c(-1, -9)]
names(train)
```

##4

```{r}
# a
glm1 <- glm(mpglevel~., data=train, family="binomial")
# b
probs1 <- predict(glm1, newdata=test, type="response")
pred1 <- ifelse(probs1>=0.5, 1, 0)
# c
table(pred1, test$mpglevel)
# d
cat("\nMean accuracy is:", mean(pred1 == test$mpglevel))
```

##5

```{r}
# a
library(class)
pred2 <- knn(train=train[,-8], test=test[,-8], cl=train[,8], k=1)
# b
table(pred2, test$mpglevel)
cat("\nMean accuracy is:", mean(pred2 == test$mpglevel))
```

##6

```{r}
# a
scale_Auto <- scale(Auto[,c(-1,-9,-10)]) # omit name, mpg, and mpglevel columns
scale_train <- scale_Auto[sampleInt, ]
scale_test <- scale_Auto[-sampleInt, ]
# b
scale_train_labels <- Auto[sampleInt, 10]
# c
scale_test_labels <- Auto[-sampleInt, 10]
# d
pred3 <- knn(train=scale_train, test=scale_test, cl=scale_train_labels)
# e
table(pred3, scale_test_labels)
cat("\nMean accuracy is:", mean(pred3==scale_test_labels))
```

##7

a. The relationship between mpg and horsepower is negative and slightly nonlinear. The relationship between mpg and is also negative and slightly nonlinear.

b. In the first two, the independent variable is a double, which could be a very wide range of different values; in the second two, the independent variable is a factor variable, which can only be one of two different values.

c.The name variable is a qualitative value with no bearing on a vehicle's performance, so using it to predict mpglevel wouldn't make sense.

d. We created the relationship between mpg and mpglevel ourselves, so it would be trivial to use mpg as one of the predictors of mpglevel.

e. The mean accuracies of models 1 and 2 are each 0.8877551, and they had an exact equal amount of predictions in each column. Our third model, which uses a scaled test set, is more accurate, and has a mean accuracy of 0.9183673.

f. A logistic regression model predicts the probabilities of outcomes, whereas a kNN with k=1 predicts just the outcomes by assigning the outcome from the closest point in the train set to a given test point.